### 2021-08-20 paper/paper.Rmd
Add paper

---
title: "Impacts of draft proposed Marine Protected Areas on population indexes derived from DFO research surveys"
author: "|
  Sean C. Anderson$^1$^[Corresponding author: sean.anderson@dfo-mpo.gc.ca], Other authors\n| $^1$Pacific Biological Station Fisheries and Oceans Canada, \n| Nanaimo, BC, Canada\n|
  $^2$Other addresses\n"
output:
    bookdown::pdf_document2:
      toc: false
      number_sections: false
      fig_caption: true
csl: mee.csl
bibliography: refs.bib
link-citations: yes
header-includes:
  \newcommand{\beginsupplement}{
  \setcounter{equation}{0}
  \renewcommand{\theequation}{S.\arabic{equation}}
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}}
editor_options:
  markdown:
    wrap: sentence
abstract: |
  Abstract here
---

```{r libs, echo = FALSE, include = FALSE, message = FALSE, warnings = FALSE}
library(here)
library(ggplot2)
library(dplyr)
library(knitr)
```

```{r options, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warnings = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

```{r load-data}

```

# Introduction

# Methods

# Results

# Discussion

# Acknowledgements

# Figures

```{r, out.width="\\textwidth", fig.cap="HBLL N geo indexes."}
include_graphics(here("figs/index-hbll-geo-restricted.pdf"))
```

```{r, out.width="\\textwidth", fig.cap="HBLL N RE."}
include_graphics(here("figs/index-hbll-geo-restricted-re.pdf"))
```

```{r, out.width="6in", fig.cap="HBLL N CV."}
include_graphics(here("figs/index-geo-hbll-cv-ratio-dotplot.pdf"))
```

```{r, out.width="6in", fig.cap="HBLL N MARE."}
include_graphics(here("figs/index-geo-hbll-mare-dotplot.pdf"))
```

```{r, out.width="\\textwidth", fig.cap="SYN geo indexes."}
include_graphics(here("figs/index-syn-geo-restricted-highlight.pdf"))
```

```{r, out.width="\\textwidth", fig.cap="SYN geo RE."}
include_graphics(here("figs/index-syn-geo-restricted-re-highlights.pdf"))
```

```{r, out.width="6in", fig.cap="SYN  CV."}
include_graphics(here("figs/index-geo-syn-cv-ratio-dotplot.pdf"))
```

```{r, out.width="6in", fig.cap="SYN MARE."}
include_graphics(here("figs/index-geo-syn-mare-dotplot.pdf"))
```

\clearpage

# References

::: {#refs}
:::

```{=tex}
\clearpage
\beginsupplement
```

# Supplemental Materials