### 2021-11-16 analysis/sim.Rmd
Add to sim.R; fork over to .Rmd

---
title: Simulating and detecting recovery in closed areas
---

Goal:

- Simulate recovery in restricted areas to assess whether current survey effort is sufficient to detect population recovery

- Approach: 
    1. fit geostatistical models to observed data
    1. use parameters from that model to simulate new data with observations at the actual historically observed locations
    1. when simulating, simulate recovery at some rate within closed areas and a stationary abundance/density outside of closed areas

- Dimensions that will likely affect the answer:
    1. species (therefore estimated spatial and spatiotemporal SD, spatial correlation range, observation error)
    1. rate of 'recovery'
    1. number of years observed
    1. whether one assesses all restricted areas together or individually

TODO:

- include depth effect? (for now, no, just a spatial random field)
- what to do with time in conditioning model? s(year)? AR1? linear? independent? (for now s(year, k = 5))
- which tests to do after to assess recovery? (a bunch of GLMs etc. are here)
- what time spans to check? (starting with all historical HBLL N; 13 years; 7 survey years)
- what possible 'recoveries' to check? (starting with 5% per year)
- fix vs. simulate spatial and/or spatiotemporal random effects? (for now fixing spatial but not spatiotemporal fields)
- all MPAs together and/or look at some individual ones? (starting with all)
- which species to work with (starting with dogfish for no good reason other than lots of data)

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  cache = TRUE,
  autodep = TRUE
)
```

```{r libs}
# remotes::install_github("pbs-assess/sdmTMB", ref="sim2")
library(sdmTMB)
library(dplyr)
library(ggplot2)
library(mgcv)
library(here)
theme_set(ggsidekick::theme_sleek())
options(dplyr.summarise.inform = FALSE)
source(here("analysis/functions.R"))
```

```{r seed}
SEED <- 2938
set.seed(SEED)
```

## Try  HBLL N dogfish

```{r data-prep-model-fit, warning=FALSE, message=FALSE}
dat_to_fit <- readRDS(here("data-generated/dat_to_fit_hbll.rds"))
grid <- readRDS(here("data-generated/hbll-n-grid-w-restr.rds"))
grid$survey_abbrev <- "HBLL OUT N"
ll_removed <- readRDS(here("data-generated/hu_co_demersalfishing_bottomlongline_d_X.rds"))
d <- filter(dat_to_fit, species_common_name == "north pacific spiny dogfish")

m0 <- fit_geo_model(d, pred_grid = grid, survey = "HBLL",
  family = "tweedie",
  # MPA_trend = TRUE,
  return_model = TRUE, do_fit = FALSE)

mesh <- make_mesh(m0$data, xy_cols = c("X", "Y"), cutoff = 15)
mesh$mesh$n
dat <- m0$data
m <- sdmTMB(response ~ 1 + s(year, k = 5), data = dat, family = tweedie(), mesh = mesh,
  silent = TRUE, time = "year", spatiotemporal = "IID")
m
```

```{r include=FALSE}
# s <- sdmTMB::sdmTMB_simulate(
#   previous_fit = m,
#   simulate_re = TRUE
# )
# 
# p <- predict(m, newdata = NULL)
# ggplot(p, aes(X, Y, colour = omega_s)) + geom_point() +
#   scale_colour_gradient2()
# 
# ggplot(s, aes(X, Y, colour = omega_s)) + geom_point() +
#   scale_colour_gradient2()
# 
# ggplot(dat, aes(X, Y, colour = response)) + geom_point() +
#   scale_color_viridis_c(trans = "sqrt")
# 
# ggplot(s, aes(X, Y, colour = observed)) + geom_point() +
#   scale_color_viridis_c(trans = "sqrt")

```

What actually happened in vs. outside restricted areas?

```{r index-real}
pred_grid <- grid
pred_grid <- filter(pred_grid, year %in% unique(dat$year))
p <- predict(m, newdata = pred_grid, sims = 200L)
p_mpa <- p[pred_grid$restricted, ]
attr(p_mpa, "time") <- "year"
p_out <- p[!pred_grid$restricted, ]
attr(p_out, "time") <- "year"
ind_mpa <- get_index_sims(p_mpa)
ind_out <- get_index_sims(p_out)
ind_real <- ind_mpa %>% mutate(type = "Restricted") %>%
  bind_rows(mutate(ind_out, type = "Not restricted"))
ind_real %>%
  group_by(type) %>%
  mutate(lwr = lwr / est[1], upr = upr / est[1], est = est / est[1]) %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr, colour = type, fill = type)) +
  geom_ribbon(alpha = 0.1, colour = NA) + geom_line() +
  scale_fill_brewer(palette = "Set1") +
  scale_colour_brewer(palette = "Set1")
```

Simulate new data:

- Keep spatial random effects as estimated (constant stuff like depth)
- Simulate new spatiotemporal random effects(?) (temperature, predator/prey etc.)
- Simulate new underlying fixed effects in BACI design
- Increase mean density within MPA by 5% per year
- Keep mean density constant within MPA
- Note the spatiotemporal random effect draws will not always have exactly mean zero
  so we can still see random variation year to year.

```{r sim-dat, warning=FALSE, message=FALSE}
b1 <- tidy(m)
b2 <- tidy(m, "ran_pars")
b <- bind_rows(b1, b2)

dat$year_covariate <- dat$year - min(dat$year)
mpa_increase_per_year <- log(1.05)

pars <- m$tmb_obj$env$last.par.best
omega_s <- pars[names(pars) == "omega_s"]
epsilon_st <- matrix(pars[names(pars) == "epsilon_st"], ncol = length(unique(dat$year)))

s <- sdmTMB::sdmTMB_simulate(
  formula = ~ 1 + restricted * year_covariate,
  data = dat,
  mesh = mesh,
  family = tweedie(),
  time = "year",
  # rho = 2 * plogis(m$model$par[['ar1_phi']]) - 1, # TODO!?
  rho = NULL,
  sigma_O = b$estimate[b$term == "sigma_O"],
  sigma_E = b$estimate[b$term == "sigma_E"],
  phi = b$estimate[b$term == "phi"],
  tweedie_p = b$estimate[b$term == "tweedie_p"],
  range = b$estimate[b$term == "range"],
  fixed_re = list(omega_s = omega_s, epsilon_st = NULL, zeta_s = NULL),
  # (Intercept), restrictedTRUE, year_covariate, restrictedTRUE:year_covariate
  B = c(b$estimate[b$term == "(Intercept)"], 0, 0, mpa_increase_per_year),
  seed = SEED
)
s <- rename(s, restricted = restrictedTRUE)
```

Real data:

```{r plot-real, fig.asp=1}
p <- predict(m)
ggplot(p, aes(X, Y, colour = omega_s)) + geom_point() +
  scale_colour_gradient2()

ggplot(p, aes(X, Y, colour = epsilon_st)) + geom_point() +
  scale_colour_gradient2() +
  facet_wrap(vars(year))

ggplot(dat, aes(X, Y, colour = response + 1)) + geom_point() +
  scale_color_viridis_c(trans = "log10") +
  facet_wrap(vars(year))
```

Simulated data:

```{r plot-sim-dat, fig.asp=1}
ggplot(s, aes(X, Y, colour = omega_s)) + geom_point() +
  scale_colour_gradient2()

ggplot(s, aes(X, Y, colour = epsilon_st)) + geom_point() +
  scale_colour_gradient2() +
  facet_wrap(vars(year))

ggplot(s, aes(X, Y, colour = log(mu))) + geom_point() +
  scale_color_viridis_c() +
  facet_wrap(~year) +
  geom_point(data = filter(s, restricted == 1L),
    pch = 21, colour = "black", fill = NA)

ggplot(s, aes(X, Y, colour = observed + 1)) + geom_point() +
  scale_color_viridis_c(trans = "log10") +
  facet_wrap(~year) +
  geom_point(data = filter(s, restricted == 1L),
    pch = 21, colour = "black", fill = NA)

s %>%
  group_by(restricted, year) %>%
  summarise(m = mean(eta)) %>%
  group_by(restricted) %>%
  mutate(m = m / m[1]) %>%
  ggplot(aes(year, m, colour = as.factor(restricted))) +
  geom_line()
```

```{r fit-sim-dat}
# s$year_covariate <- s$year_covariate - mean(s$year_covariate)
m2 <- sdmTMB(
  observed ~ 1 + restricted * year_covariate,
  data = s,
  mesh = mesh,
  time = "year",
  # spatiotemporal = "AR1",
  spatiotemporal = "IID",
  family = tweedie(),
  silent = TRUE
)

pred_grid <- grid
pred_grid <- filter(pred_grid, year %in% unique(dat$year))
pred_grid$year_covariate <- pred_grid$year - min(pred_grid$year)
p <- predict(m2, newdata = pred_grid, sims = 200L)

p_mpa <- p[pred_grid$restricted, ]
attr(p_mpa, "time") <- "year"

p_out <- p[!pred_grid$restricted, ]
attr(p_out, "time") <- "year"

ind_mpa <- get_index_sims(p_mpa)
ind_out <- get_index_sims(p_out)

ind <- ind_mpa %>% mutate(type = "Restricted") %>%
  bind_rows(mutate(ind_out, type = "Not restricted"))
```

## How to assess improvement in closed areas combined?

```{r assess-sim-dat}
# 1. visual on geostat index?
ind %>%
  group_by(type) %>%
  mutate(lwr = lwr / est[1], upr = upr / est[1], est = est / est[1]) %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr, colour = type, fill = type)) +
  geom_ribbon(alpha = 0.1, colour = NA) + geom_line() +
  scale_fill_brewer(palette = "Set1") +
  scale_colour_brewer(palette = "Set1")

# 2. fancy geostat BACI?
tidy(m2)

# 3. naive GLM BACI on geostat output?
# glm(est ~ year, data = ind_mpa, family = Gamma(link = "log"))
# glm(est ~ year, data = ind_out, family = Gamma(link = "log"))
a <- glm(est ~ type*year, data = ind, family = Gamma(link = "log"))
broom::tidy(a)
# above ignores autocorrelation

# 3.1 weight by 1/SE?
a <- glm(est ~ type*year, data = ind, family = Gamma(link = "log"), weights = 1/se)
broom::tidy(a)

# 3.2 via lm()?
a <- lm(log(est) ~ type*year, data = ind)
broom::tidy(a)

# 3.3 via gls() with AR1 residuals?
a <- gls(log(est) ~ type*year, data = ind, correlation = corAR1(form = ~ 1 | type))
summary(a)

# 4. naive GLM BACI on raw observations?
a1 <- mgcv::gam(observed ~ restricted * year_covariate, data = s, family = tw())
# above ignores spatial correlation and temporal autocorrelation and sampling location
summary(a1)

# 5. plot of mean by area?
s %>%
  group_by(restricted, year) %>%
  summarise(m = mean(eta)) %>%
  group_by(restricted) %>%
  mutate(m = m / m[1]) %>%
  ggplot(aes(year, m, colour = as.factor(restricted))) +
  geom_line()
```