### 2023-12-05 text/main.tex
Revisions

% !TEX TS-program = pdflatexmk

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1.15in]{geometry}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage[dvipsnames]{xcolor}
\definecolor{niceblue}{HTML}{236899}
\newcommand{\rev}[1]{{\color{niceblue} #1}}
\newcommand{\R}[1]{\label{#1}\linelabel{#1}}
\newcommand{\lr}[1]{line~\lineref{#1}}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=black,
    urlcolor=niceblue,
    citecolor=niceblue,
    linkbordercolor = white
}
\usepackage{longtable}
\usepackage{amsbsy}
\usepackage{epsfig}
\usepackage{bm}
\usepackage{xspace}
\usepackage{color}
\usepackage{lineno}
\usepackage{ragged2e}
\usepackage{comment}

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
\useosf % osf for texb, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}
\usepackage{setspace}

\makeatletter
\let\LN@align\align
\let\LN@endalign\endalign
\renewcommand{\align}{\linenomath\LN@align}
\renewcommand{\endalign}{\LN@endalign\endlinenomath}
\let\LN@gather\gather
\let\LN@endgather\endgather
\renewcommand{\gather}{\linenomath\LN@gather}
\renewcommand{\endgather}{\LN@endgather\endlinenomath}
\makeatother

\widowpenalty10000
\clubpenalty10000

\definecolor{darkgrey}{HTML}{A9A9A9}
\renewcommand\linenumberfont{\normalfont\bfseries\small\color{darkgrey}}
\modulolinenumbers[2]

\usepackage{booktabs}
\usepackage[round]{natbib}
\bibliographystyle{icesjms}
\bibpunct{(}{)}{,}{a}{}{,}
\usepackage{authblk}
\usepackage{pdflscape}
\usepackage{setspace}

\newcommand{\pe}[1]{{\color{blue}#1}}
\newcommand{\sa}[1]{{\color{red}#1}}
\newcommand{\kg}[1]{{\color{violet}#1}}
% other colours: orange, magenta, sky blue, purple, grey

\newcommand*{\TitleFont}{%
      \usefont{\encodingdefault}{\rmdefault}{b}{n}%
      \fontsize{13}{15}%
      \selectfont}

% \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
% \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
% \newcommand{\lr}[1]{line~\lineref{#1}}

\date{}

\title{Impacts on population indices if scientific surveys are excluded from marine protected areas}

% Tentative order (alphabetical after Philina right now) ...
\author[1,2]{Sean C. Anderson}
\author[1]{Philina A. English}
\author[3]{Katie S.P. Gale}
\author[1,4]{Dana R. Haggarty}
\author[5]{Carolyn K. Robb}
\author[3,6]{Emily M. Rubidge}
\author[3,7]{Patrick L. Thompson}

\affil[1]{Pacific Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada}
\affil[2]{Department of Mathematics, Simon Fraser University, Burnaby, BC, Canada}
\affil[3]{Institute of Ocean Sciences, Fisheries and Oceans Canada, Sidney, BC, Canada}
\affil[4]{Department of Biology, University of Victoria, Victoria, BC, Canada}
\affil[5]{Regional Headquarters, Fisheries and Oceans Canada, Vancouver, BC, Canada}
\affil[6]{Department of Forest and Conservation Sciences, University of British Columbia, Vancouver, BC, Canada}
\affil[7]{Department of Zoology, University of British Columbia, Vancouver, BC, Canada}
\affil[*]{corresponding author: sean.anderson@dfo-mpo.gc.ca}

\input{values} % R output

% \doublespacing
% \onehalfspacing
\begin{document}
\begin{spacing}{1.4}
\maketitle

% \RaggedRight

\clearpage
\linenumbers

\section*{Abstract}

% ICES: 200 words max
% perhaps could frame this as broader than MPAs (e.g., mostly wind farms in the US) and emphasize the general approach more with the NSB MPA case study

Marine protected areas (MPAs) are increasingly common worldwide, typically restricting fishing activities. However, MPAs may also limit scientific surveys that impact benthic habitat. We develop an approach that combines historical data degradation and simulation to investigate the effects of excluding surveys from MPAs on population abundance indices. Our approach quantifies losses in precision, inter-annual accuracy, trend accuracy, and the power to detect trends, as well as correlates of these effects. We apply this approach to a proposed MPA network off western Canada, examining \nSpp\ groundfish species observed by four surveys. Survey exclusion particularly impacted less precise indices, species well-represented in MPAs, and those whose density shifted in or out of MPAs. Redistributing survey effort outside MPAs improved precision but had limited benefits for accuracy or trend detection---it \R{B17}can make our estimates more precise about the ``wrong'' index. While these changes may not qualitatively alter stock assessment for many species, in some cases $\approx$30 percentage point reductions in the power to detect simulated 50\% population declines suggest meaningful impacts are possible. If survey restrictions continue expanding, index integrity could further degrade, eventually compromising the management of exploited populations. Regulating surveys within MPA boundaries therefore requires careful consideration to balance MPA objectives with the need for reliable data.

\vspace{5mm}
\noindent
Keywords:
groundfish,
index of abundance,
MPA,
spatiotemporal,
survey effort reduction,
scientific survey exclusion

% \clearpage
% \tableofcontents
% \thispagestyle{empty}
% \setcounter{page}{0}
% \clearpage

% ER: FYI Protection Standards to better conserve our oceans (dfo-mpo.gc.ca)
% These minimum standards to MPAs apply to federal MPAs only but may be adopted by other policy tools, and include the following gear restrictions:
% The new protection standard for MPAs prohibit four key industrial activities in all new federal MPAs: oil and gas activities; mining; dumping; and bottom trawling.
% The prohibition on bottom trawling applies to mobile bottom-contact gear used for commercial and recreational purposes, including otter trawls, beam trawls, shrimp trawls, hydraulic clam dredges, and scallop dredges.
% All other activities, including fishing that does not use mobile bottom-contact gear, will continue to be assessed on a case-by-case basis to ensure they do not pose a risk to the conservation objectives of the MPA.
% *Bottom trawling for Indigenous food, social, and ceremonial purposes and for scientific research purposes will be allowed within MPA where it does not pose a significant risk to the MPA's conservation objectives.*
%
% This last point is important because it clearly states that science surveys can continue if no risk to cons. Obj.  of MPA (unlikely for most of our MPAs as benthic features often Conservation Objectives).

\section*{Introduction}

% <!-- Paragraph 1: scientific surveys are important and pervasive -->
Scientific surveys form the backbone of population status assessment for exploited marine species \citep{doubleday1981, hilbornwalters1992, gunderson1993}. One of the primary products of such surveys are indices of relative or absolute stock biomass or abundance through time (herein population or stock ``indices''). These indices are vital to most fisheries stock assessment \citep{hilbornwalters1992} and to other applications such as the assessment of species at risk \citep{iucn2012}. Because of this importance, government agencies spend considerable effort to maintain random, stratified random, or fixed-station sampling designs, as well as consistent gear and survey timing, so that indices remain proportional to underlying trends in population biomass or abundance.

% <!-- Paragraph 2: globally MPAs are being proposed more and more -->
Globally, regions of the ocean are increasingly set aside as reserves or marine protected areas (MPAs) to protect ecosystems, species groups, fisheries, and culturally and socially valued areas from a range of anthropogenic threats. Motivated in part by the direction of the United Nations Convention on Biological Diversity to protect at least 10\% of marine areas by 2020 (since increased to 30\% by 2030), the area of MPAs worldwide has increased from 0.7\% to $\sim$8\% over the last 22 years \citep{WDPA2022}. Although the level of protection of individual MPAs varies from highly protected to multiuse \citep{hortaecosta2016, grorud-colvert2021}, it is clear that full- or high-protection MPAs (i.e., low or no extractive or destructive activities) have the best conservation outcomes \citep[e.g.,][]{lester2008, sciberras2015, edgar2014, zupan2018}. Commercial bottom trawling is therefore frequently restricted in MPAs. In Canada, for example, excluding industrial bottom trawling has been recommended as a ``Protection Standard'' for all conservation areas that meet the IUCN definition of an MPA \citep{dfo2022standards}.

% <!-- Paragraph 3: restricted from fishing, but then the question becomes scientific surveys -->
While MPA regulations typically target commercial and recreational fishing activities, a focus on building more effective MPAs has also raised concerns about impacts on benthic habitat from scientific survey gear \citep{field2006, ursgroup2016, saarman2018, benoit2020national}. Although research surveys typically have much lower benthic footprints than commercial fleets---primarily because of their considerably smaller scale \citep{benoit2020national}---hook-and-line survey gear, and in particular trawl gear, damages benthic habitat \citep[e.g.,][]{collie2000a, kaiser2006, hiddink2017}. Scientific research surveys that pose a risk to conservation objectives may be excluded from MPAs and fisheries closures. In Canada, for example, a permitting process usually determines if research activities constitute an exception under MPA regulations. For example, Rockfish Conservation Areas, a collection of over 160 areas in Pacific Canada aiming to protect sensitive rockfish populations, prohibit several modes of commercial and recreational and fishing as well as trawl or longline extractive research surveys \citep{thornborough2020}.

% <!-- # Paragraph 4: if excluded, this presents a number of possible issues -->
If scientific surveys are excluded from MPAs based on the gear used and negative impacts to benthic conservation objectives, there are several concerns related to the precision, inter-annual accuracy, and bias in indices derived from these surveys \citep{field2006, rideout2017, saarman2018, benoit2020national, benoit2020gulf, ices2020, ices2023}.
A reduced number of survey sets is expected to reduce index precision according to statistical sampling theory \citep{cochran1977}, although survey effort may be redirected to the remaining survey domain.
If relative abundance inside and outside the closed areas changes because of fish movement or recovery within closed areas, a loss of survey coverage from portions of a population's distribution may also introduce bias \citep[][]{benoit2020national}.
An analyst faced with a restricted survey must furthermore decide whether to ``shrink'' the survey domain by removing the restricted area \citep{rideout2017, benoit2020national} or extrapolate/interpolate from the surveyed region into the restricted area \citep{benoit2020national}.
\R{A2}This decision can also impact index quality and therefore scientific advice to management (e.g., stock status and catch recommendations), for example by dissociating the survey domain from the stock domain in the former case or by imposing survey patterns in the restricted area based on surrounding observations in the latter case.

% <!-- # Paragraph 5: what we do here -->
Here, we develop a general approach for evaluating the impacts of restricting scientific surveys on population indices through a retrospective analysis of survey data.
We apply this approach to a proposed MPA network off western Canada---the Northern Shelf Bioregion MPA Network \citep[][Fig.~\ref{fig:map}]{dfo2022networkactionplan}.
We use spatiotemporal-model-based and traditional design-based methods to derive indices of relative biomass or abundance for four surveys (three using trawl gear and one using long-line gear) and \nSpp\ groundfish species and recalculate those indices as if scientific surveys had been restricted from areas proposed in the MPA network.
We evaluate changes in index precision, accuracy, trend bias, and power to detect decline, and evaluate correlates of these changes.

\section*{Methods}

We first describe a general approach for evaluating the impact of proposed survey restriction on indices of abundance or biomass with existing data.
We then describe the specifics of how we applied this approach to four groundfish surveys within a proposed MPA network off the coast of western Canada.

\subsection*{Calculating survey indices}

We derived both design-based and spatiotemporal-model-based (herein ``model-based'') indices of abundance from survey data.
Our results, however, focus on model-based indices because of a general move towards this approach, because the approach tends to have better precision than design-based indices \citep{thorson2015a}, and because model-based approaches are likely better suited to dealing with unanticipated changes to survey design \citep{ices2023}.

Our design-based random stratified estimation follows \citet[][p.~91]{cochran1977} and takes the mean abundance within each depth stratum and the mean of those means.
We used a bootstrap procedure to quantify variance of this estimator \citep{schnute2000}.
The bootstrap procedure performed the same index calculation after sampling with replacement by depth stratum 1000 times and calculating the standard deviation of the resulting means \citep{schnute2000}.
We also calculated a classic stratified random sampling variance \citep[][p.~95]{cochran1977} although the results were similar to the bootstrap procedure and so we do not include them here.

We derived model-based indices of relative abundance or biomass using a geostatistical modelling approach with spatial and spatiotemporal Gaussian Markov random fields describing latent spatial and spatiotemporal processes \citep[e.g.,][]{thorson2015a, anderson2022}.
Depending on the species and survey, these models were configured as either delta models \citep{aitchison1955} with a logit link for a binomial model of encounter probability  and log link for a gamma model of positive catch weight \citep{schnute2003} or as \citet{tweedie1984} observation models with a log link for catch weight.
We included an offset term \citep[][p.~206]{mccullagh1989} for log area swept for the trawl models and log hook count \R{B10} for the long-line models.
Following a common model configuration for index standardization \citep[e.g.,][]{thorson2015a, thorson2019a}, our models included an independent intercept each year, a fixed spatial random field, and independent spatiotemporal random fields.
We reduced model complexity as needed based on diagnostics to reflect an analyst aiming for adequate model convergence and parsimony (Table~\ref{tab:model-configs}, Supporting Material).
We fit our models with the sdmTMB \textsf{R} package \citep{anderson2022}, which implements the models through TMB \citep{kristensen2016} with input matrices generated through \textsf{R}-INLA \citep{lindgren2015}.
Additional modelling details are available in the Supporting Material.

% \subsection*{Comparing indices to the status quo}

We calculated a series of indices. The first two form the main analysis and the third is an additional option we focus on in supplemental figures.

\begin{enumerate}

\item Status quo: estimate the index based on all available data and the full survey domain. This becomes our comparison index when assessing changes to precision, accuracy, or bias.

\item Shrunk: estimate the index based on survey observations outside the proposed MPA network and a reduced survey domain that excludes the MPA network. We expect this to be the most likely outcome if survey effort is restricted.

\item Extrapolated: using observations outside the MPA network, estimate the area-weighted index of abundance for the full status quo survey domain. This option could take advantage of nearby data and spatial correlation to predict density within the MPA and retain the full survey domain.
\end{enumerate}

\subsection*{Metrics of precision, accuracy, and trend bias}

We compared performance measures of the restricted shrunk or extrapolated indices to the status quo indices that use the full set of historical survey observations:

\begin{enumerate}
    \item Precision: We calculated the annual coefficient of variation $\mathrm{CV}_{t}$ for year $t$ as $\sqrt{\exp(\mathrm{SE}_{t}^2) - 1}$, where $\mathrm{SE}$ is the standard error on the index as calculated with the generalized delta method in TMB \citep{kristensen2016} for the model-based index. We then calculated the proportional change in $\mathrm{CV}_{t}$ between each restricted index and the status quo index. This represents a standardized measure of precision loss.

    \item Accuracy: We calculated the annual relative error (RE) as $(\hat{\theta}_t - \theta_t) / \theta_t$, where $\hat{\theta}_t$ represents the index estimate from the restricted model in year $t$ and $\theta_t$ represents the index estimate from the status quo survey. We first centered $\hat{\theta}_t$ and $\theta_t$ by dividing them by their respective geometric means across years to avoid including consistent bias (an index from a shrunk survey domain should be lower than a full-survey domain index) and because surveys are often considered relative indices of abundance or biomass with an estimated catchability parameter that is estimated in stock assessments. We then calculated the absolute RE. The median absolute RE (MARE) represents a measure of accuracy loss---the expected magnitude of relative error in a typical year.

    \item Trend bias: We determined the trend in the relative error described above by fitting a linear regression to RE by year and reporting the slope in units of RE per decade.

\end{enumerate}

\subsection*{Correlates of survey-restriction impacts}

We examined three potential predictors of changes in precision, accuracy, and trend bias:
(1) the CV of the status quo index, (2) the biomass or abundance proportion of a species inside the MPA network, and (3) trends in proportion of a species inside vs.\ outside the MPA.
When modelling MARE as the response, we fit generalized linear models (GLMs) with gamma observation error and log links with log-transformed predictor variables to identify patterns.
This implies that a proportional change in the predictor corresponds to a proportional change in the response.
% For any species with slight proportional decreases in precision, we set these values to small positive values (0.01) for the purposes of this GLM because the log link requires positive data.
When modelling percent increase in CV as the response (which had some negative values), we fit a model with both the mean and observation variance growing linearly with the log-transformed predictor (log proportion stock in MPA).
% TODO loss)!

TODO!!!!!!!!!!!!!!!!!!!!!!!!!!! FIX LOG PREDICTOR?? WHAT DID I DO !!!!????????????????

For the third, we compared the slope of relative error per decade against slope of density proportion within the MPA per decade for the most impacted survey(s) and fit a linear regression.
% We did not fit these models to evaluate statistical significance (noting that we ignore uncertainty in the predictor and response variables) but to characterize the relationships.
Because our analysis applies to proposed MPAs, the MPAs should not have had any causal effect on historical biomass or abundance.

\subsection*{Examining the role of effort reduction}

If surveys are restricted from MPAs, a likely scenario is that survey effort would be redistributed throughout the remaining survey domain.
We therefore examined two additional scenarios to assess the contribution of effort reduction itself vs.\ the specific spatially blocked nature imposed by the MPAs.
(1) Down-sampled: do not remove survey sets from the MPAs explicitly, but remove an equivalent level of survey effort by randomly removing survey sets from the entire survey domain, with equal numbers removed from each survey depth stratum.
(2) Up-sampled: redistribute the lost survey effort by assigning an equivalent number of new survey sets within each stratum and generate plausible observations for these new locations by simulating new observations from the status quo spatiotemporal model (Supporting Methods).
We created five random versions of each resampled dataset, calculated the same indices as the main analysis, and calculated the same metrics of precision, accuracy, and trend bias.

% TODO: debatably, this should come a bit earlier:

\subsection*{Power analysis of the reduced ability to detect decline}

We performed a power analysis to identify the reduced ability to detect an exponential population decline under the MPA-restricted scenario compared to the status quo.
We used the spatiotemporal models fit to the full dataset (Supporting Material) and simulated new datasets with the fitted spatial random fields, new simulated spatiotemporal random fields, and observation error based on the fitted distribution.
In the simulation, we replaced the year factor effects with exponential declines in biomass or abundance density as a linear function in log link space.
For models simulated from a delta-gamma distribution, we held the detection probability constant.
We chose rates of decline that would be meaningful \R{A4}in a conservation context and that illustrated contrast between the two sampling scenarios: 50\% decline for species with average status quo CVs of $>$ 0.25 and 30\% decline for species with CVs $\le$ 0.25 over the span of the surveys (15 to 18 years).

We then fit spatiotemporal models (omitting spatiotemporal random effects for simplicity, consistency, and estimation speed) with the same observation family and tested whether the upper 95\% confidence interval (CI) on the slope parameter representing the effect of year was less than zero.
We repeated the simulation and estimation 300 times to calculate power: the proportion of times that the estimation models correctly identified a negative population trend.

\subsection*{Application to an MPA network off the west coast of Canada}

\subsubsection*{Proposed Northern Shelf Bioregion MPA network}

We applied the above approach to an MPA network and set of surveys off the west coast of Canada.
The MPA network planning area, the Northern Shelf Bioregion (NSB), is approximately 102,000~km$^2$ and the proposed MPA network encompasses about 30\% of this area (30,493~km$^2$).
% TODO: this still correct?
The network process has been co-led by 15 partner First Nations, the province of British Columbia, and the Canadian federal government.
% (Fisheries and Oceans Canada, Environment and Climate Change Canada, Parks Canada, Natural Resources Canada, Transport Canada).
% CR: Parks Canada, EEEC and DFO? Might be worth mentioning, just because they represent a diversity of spaces (not just fisheries management).
% TODO: ER: is this correct?
Following a systematic conservation planning approach \citep{margules2000}, its design was guided by national and regional goals \citep{canada2011, canada2014} and centered on meeting network conservation objectives \citep{dfo2022networkactionplan}.
Of the six goals of the network, Goal 1 is of primary importance: ``protect and maintain marine biodiversity, ecological representation and special natural features.''
Additional goals and objectives include the protection of traditional use and cultural heritage, contributing to the protection of fisheries resources and their habitat, and providing opportunities for scientific research, education, and awareness \citep{canada2014}.
The proposed network is built off existing areas, where nearly 50\% of the draft scenario footprint already exists as some type of spatial protection (existing provincial or federal MPAs or fisheries closures in the form of Rockfish Conservation Areas) leaving $\sim$50\% of the network as new areas identified through the planning process \citep{dfo2022networkactionplan}.
The proposed NSB MPA network was announced by the planning partners in February 2023 with plans for iterative implementation; the first set of new MPAs are expected to be established by 2025.
Our analysis makes some necessary assumptions about what areas of the network may be restricted from survey activity (Supporting Material: Northern Shelf Bioregion zones).

\subsubsection*{Scientific surveys}

We analyzed data from four research surveys including three synoptic bottom trawl surveys \citep[][Fig.~\ref{fig:map}]{sinclair2003syn, anderson2019synopsis}:
% Queen Charlotte Sound \citep[SYN QCS,][]{williams2018synqcs},
% Hecate Strait \citep[SYN HS,][]{wyeth2018synhs}, and
% West Coast Haida Gwaii \citep[SYN WCHG,][]{williams2018synwchg}.
Queen Charlotte Sound (SYN QCS),
Hecate Strait (SYN HS), and
West Coast Haida Gwaii (SYN WCHG).
% We omitted two related surveys off the west coast of Vancouver Island because they are not affected by the proposed MPA network.
The included surveys cover the continental shelf and upper slope off the northern part of the west coast of Canada (Fig.~\ref{fig:map}) and primarily target soft-bottom habitat. They have been conducted since 2003, are typically run in alternating years, and follow a random depth-stratified design with sampling units of 2 km $\times$ 2 km blocks with consistent gear and fishing protocols \citep{sinclair2003syn}. The unit of measurement for the synoptic trawl surveys is fish biomass per area swept.

The fourth survey we analyzed was the Outside Northern Hard Bottom Longline survey (HBLL OUT N), which covers a similar geographic region to the synoptic trawl surveys, but instead focuses on hard bottom habitat \citep{doherty2019hbllout} (Fig.~\ref{fig:map}). This survey has generally been run in alternating years since 2006, is conducted by Fisheries and Oceans Canada in collaboration with the Pacific Halibut Management Association, and also follows a random depth-stratified design with sampling units of 2 km $\times$ 2 km blocks \citep{doherty2019hbllout, anderson2019synopsis}. The unit of measurement for the HBLL OUT N survey is numbers of fish per assumed area of catchability based on 8 ft (2.438 m) hook spacing and an assumed catch radius of 30 ft (9.144 m).

Based on a threshold of at least 5\% of survey sets detecting a species, we selected \nSpp\ species of benthic fishes including rockfishes (\textit{Sebastes}), flatfishes (\textit{Pleuronectiformes}), Pacific cod (\textit{Gadus macrocephalus}), and several elasmobranch species (Table~\ref{tab:spp-sci}). Of these species, \synNSpp\ met this threshold in the synoptic trawl surveys and \hbllNSpp\ met this threshold in the longline survey.

\section*{Results}

Full exclusion of scientific surveys from the proposed NSB MPA network zones would have resulted in a proportional survey domain loss of \lostHBLL, \lostWCHG, \lostHS, and \lostQCS\ across the HBLL OUT N, SYN WCHG, SYN HS, and SYN QCS surveys, respectively (Fig.~\ref{fig:map}, Fig.~\ref{fig:strata}).
% If existing MPAs also excluded survey effort, \lostHBLLall, \lostWCHGall, and \lostQCSHSall\ of the survey domain would be excluded.
The impact of the proposed MPAs on indices was variable across species and surveys when looking at the time series (Fig.~\ref{fig:timeseries}, Fig.~\ref{fig:ts-hbll}--\ref{fig:ts-qcs}).
Overall trends were often similar between the status quo and shrunk scenarios, although the shrunk scenario had larger confidence intervals and increased interannual variability.
The increased interannual variability led to smoother generalized additive model fits in some cases---flattening some patterns that were apparent with the status quo indices (e.g.,~Fig.~\ref{fig:timeseries}).

\subsection*{Metrics of precision, accuracy, and trend bias}

Restricting a survey nearly universally reduced precision (Fig.~\ref{fig:dotplot}a--d) and accuracy (Fig.~\ref{fig:dotplot}e--h) with respect to the status quo index and sometimes resulted in trend bias (Fig.~\ref{fig:dotplot}i--l).
Impacts on precision, accuracy, and trend bias were strongest in the SYN WCHG trawl survey (Fig.~\ref{fig:dotplot}a,~e,~i) followed by the HBLL~OUT~N survey (Fig.~\ref{fig:dotplot}b,~f,~j).
These two surveys contained species with the highest proportion of density within the proposed MPA network with proportions up to $\approx$0.5 (Fig.~\ref{fig:dotplot}).
For species with at least 15\% of their biomass or abundance density within the MPA, the average CV increase was \precisionWCHGshrunk\% in SYN~WCHG and \precisionHBLLshrunk\% in HBLL~OUT~N (Fig.~\ref{fig:dotplot}a, b).
Accuracy loss, MARE---reflecting proportional error in a typical bad year---ranged up to \canaryrockfishSYNWCHGslopere\ (Canary Rockfish) in SYN WCHG and \southernrocksoleHBLLslopere\ (Southern Rock Sole) in HBLL~OUT~N with an average for species with at least 15\% density within the MPA of \accuracyWCHGshrunk\ and \accuracyHBLLshrunk\ in these surveys, respectively (Fig.~\ref{fig:dotplot}e,~f).

Trends in RE ranged up to \canaryrockfishSYNWCHGslopere\ per decade for canary rockfish in SYN WCHG, meaning the index diverged from the status quo index by up to \canaryrockfishSYNWCHGsloperePerc\% per decade (Fig.~\ref{fig:dotplot}i).
In the HBLL~OUT~N, trends in RE per decade were largest for North Pacific Spiny Dogfish at \northpacificspinydogfishSYNWCHGslopere\ per decade.
Effects within the SYN~QCS, where only four species had $>$15\% density within the MPA, were the smallest (Fig.~\ref{fig:dotplot}d,~h,~l).

\subsection*{Correlates of survey-restriction impacts}

There were strong relationships between some covariates and our metrics of precision, accuracy, and trend bias for the shrunk-survey-domain indices (Figs.~\ref{fig:covariates} and ~\ref{fig:slopes}).
Indices with larger CVs tended to have greater losses of accuracy with survey restriction than indices starting with smaller CVs (Fig.~\ref{fig:covariates}a).
Across all surveys and species, this was at a rate of \covCvmaregeostat\ increase in MARE per 1\% increase in status quo index CV.
Species with a greater proportion of their density within the MPAs suffered larger losses of accuracy and precision (Fig.~\ref{fig:covariates}b, c).
Across all surveys, a 1\% increase in proportion of a species within the MPA network resulted in an increase in MARE of \covMareGeo\ and an increase in CV of \covPrecisionGeo.

There was a nearly one-to-one negative relationship between the slope of density proportion within the MPAs per decade and the slope of relative error per decade (regression slope: \REslopeRegress, Fig.~\ref{fig:slopes}).
Visualizations of the most extreme cases illustrate these trends in proportion (Fig.~\ref{fig:slopes}, right column)---Canary Rockfish, North Pacific Spiny Dogfish, Redstripe Rockfish, and Petrale Sole were increasingly found in higher density proportions outside the proposed MPA zones.
In other words, proportionally, these species were increasingly seen by the survey resulting in index time series in the restricted scenario that trended up more positively (or not down negatively enough) compared to the status quo full-survey-domain index.

Extrapolating the species distributions into the closed areas instead of restricting the survey domain reduced precision with little or no overall benefit to accuracy or trend on average (Fig.~\ref{fig:dotplot-extrapolate}).
Calculating the indices with design-based methods instead of spatiotemporal models resulted in similar qualitative conclusions (Fig.~\ref{fig:metrics-design-geo}, \ref{fig:covariates-design}) but with some minor differences.
\R{B16}The impact on accuracy loss and trend bias with increasing proportion of the stock excluded by the MPA was slightly greater for the design-based vs.\ geostatistical indices (Fig.~\ref{fig:covariates-design}).
However, while the geostatistical indices overall had lower CVs than the design-based indices with or without MPA restrictions (Fig.~\ref{fig:design-model-cv}), the geostatistical indices were proportionally more affected by survey restriction than the design-based indices (Fig.~\ref{fig:metrics-design-geo}).

\subsection*{Examining the role of effort reduction}

Through two resampling scenarios, we evaluated the contribution of reduced sampling effort to the above results (Fig.~\ref{fig:sampling}).
In the scenario where we removed an equivalent effort randomly from each strata, we observed similar precision loss to the MPA exclusion scenario but with improved confidence interval coverage, accuracy loss, and trend bias (Fig.~\ref{fig:sampling}, orange vs.\ black).
In the scenario where we up-sampled survey effort to maintain an equivalent total effort distributed across the now restricted surveyed domain,
precision loss was almost entirely ameliorated but at the expense of accuracy and poorer confidence interval coverage (Fig.~\ref{fig:sampling}, blue vs.\ black).
% TODO: CR: The implication that you make here also implies that the MPAs are biased towards some species? Did you test if it was wrong about the trend as well? In the next paragraph you look at the precision relative to the trend, but if you design calculate/estimate the model inside the MPA areas, do you get the same trend as in the "open" areas? This speaks to an analysis from Cordue 2006 who looked at untrawlable areas and concluded if the trends were going in opposite directions, that's bad, but if the trends go in the same direction its OK, cause you just fit a lower Q, precision notwithstanding (I'm paraphrasing and trying to remember the exact result). This comment you can choose to address or not, I'm mostly just curious. I guess to answer this you would fit the model in the closed area, fit it in the open area and see if the trend is the same. I guess you kind of get this from teh RE on the trend over decades.
This implies that the redistributed additional survey effort resulted in an index that was increasingly precise about the ``wrong'' values (with correct defined as the full status quo index).

\subsection*{Power analysis of the reduced ability to detect decline}

Reductions in power to detect population declines were generally stronger in the SYN WCHG trawl survey than the HBLL OUT N longline survey (Fig.~\ref{fig:power}).
As examples, in the SYN WCHG trawl survey, the probability of detecting a 50\% decline for Redstripe Rockfish based on a 95\%~CI declined from 0.64 to 0.38.
Power to detect a 30\% decline for Redbanded Rockfish declined from 0.82 to 0.55.
% See 05-power.R 'NUMBERS FOR PAPER' to grab these hardcoded values
Reduction in power was generally 10 percentage points or smaller for the HBLL OUT N survey with the largest effects seen for Copper Rockfish and Tiger Rockfish.
Some species, such as Lingcod, saw minimal effects on power for both surveys.

\section*{Discussion}

We undertook an analysis of historical survey data to assess expected impacts on population indices if scientific surveys are excluded from a proposed MPA network.
Our analysis used the NSB MPA network on the west coast of Canada as a case study, but our approach is applicable to decisions about scientific survey exclusion worldwide.
Our results documented a range of impacts on precision, accuracy, and bias in trend estimation.
Species with less precise indices, species with a higher relative density in the restricted area, and species whose distribution changed with respect to the excluded area were most impacted.
For many species, the impacts were small to moderate and are unlikely to have strong effects on science advice.
However, in other cases, impacts on precision and accuracy had notable effects on the power to detect population decline.
Furthermore, biases in population trend---driven largely by changes in distribution inside vs.\ outside the excluded areas---could impact stock assessment results and estimates of decline for conservation agencies.
Analyses such as ours are an important part of assembling the evidence needed when making decisions about restricting or allowing scientific surveys within a previously established survey domain.
Here, we discuss the magnitude of these effects, why some effects may have a greater impact than others, consequences beyond stock assessment, caveats to our approach, and make recommendations based on our analyses.

\subsection*{Are these impacts enough to make a difference?}

Our results quantified effects on precision and accuracy of survey indices.
Are these effects large enough to impact stock assessment?
For SYN QCS and SYN HS, the impacts of these survey exclusions are probably small enough that they are unlikely to have major consequences for stock assessment outcomes.
However, there may be impacts for some species within the SYN WCHG and the HBLL OUT N survey.
Our analysis found notable reductions in the power to detect 30\% and 50\% declines, which hint that these effects may meaningfully impact stock assessment estimates.
Ultimately, the impact on stock assessments will require further analysis---either with retrospective analyses degrading stock assessment input data, similar to that performed here, or through simulation.
A complete analysis would include closed-loop simulation \citep{punt2016} under restricted survey scenarios to evaluate how these effects might impact the probability of achieving conservation and fishery objectives.

% Paragraph about how surveys tend to be designed to be right at the border of
% utility to balance precision with costs and resources
Some of the less-precise survey indices may already be pushing the bounds of utility; however, they remain the most important source of information for assessing commercially fished stocks.
In the surveys examined here, precision for many stocks is already above the targeted \citep{sinclair2003syn} 0.2 CV and stocks with higher CVs were proportionally more affected by lost survey coverage.
Surveys are often designed to have just enough precision for their intended use due to the substantial resources needed to run them \citep{hilbornwalters1992}.
Furthermore, survey indices are often assumed to suffer from additional process error \R{E3a} (e.g., due to variation in spatial and vertical availability to surveys, weather and environmental conditions, gear changes, different fishing captains) and so additional error is often added when entering the index into an assessment \citep{pennington1994, francis2003}.
\R{E3b}Survey exclusion from an MPA can be thought of as adding to this additional process error since it exacerbates the effect of variation in spatial availability given the full stock domain is not accessible to the survey.
This further reduces an already imprecise index's ability to inform an assessment model.
There is no hard cut off of precision and accuracy in terms of impact on stock assessment---the impacts occur on a continuous spectrum.
Even if the specific effects identified here only have moderate impacts on stock assessment, the issue could be ``death by a 1000 cuts'' if survey restriction continues to grow in the future \citep{benoit2020national}.

We hypothesize that trend bias (trend in relative error) will have the greatest impact on stock assessment, recovery monitoring, and conservation agency assessment.
For some species, this trend bias was considerable.
For example, North Pacific Spiny Dogfish, Canary Rockfish, Redstripe Rockfish, and Petrale Sole had increases in relative error of $\approx$20--30\% per decade.
Index precision is sometimes ignored in simple stock assessments \citep{free2020} or adjusted through re-weighting schemes \citep{francis2011} and many index-based management procedures ignore precision in their calculations all together \citep{carruthers2016}.
A bias in trend, however, will directly impact index-based management procedures and affect the constant proportionality assumption of relative indices, similar to the impact of other non-trawlable habitat when species distributions change \citep{cordue2007}.
Additionally, if MPAs operate as hoped, the proportion of a species occurring inside the MPAs will initially increase over time, which would be expected to cause the restricted indices to underestimate the true overall species abundance resulting in a negative trend in relative error.
Therefore, trend bias may be even more present with real MPAs compared to our retrospective analysis where future MPAs had no causal mechanism to affect the past.
Fortunately, this potential for underestimation of increases in abundance is consistent with the precautionary principle.

We evaluated impacts on indices of abundance if surveys are excluded from closed areas; however, survey exclusion would also impact biological samples such as \R{E4}age and length composition data.
The loss of such sampling within closed areas might be expected to result in larger impacts on stock assessment than degraded indices of abundance themselves (\citeauthor{yin2004} \citeyear{yin2004}, but see \citeauthor{chen2003} \citeyear{chen2003}).
Composition data, and especially age composition data, are important inputs to modern integrated age structured stock assessments and provide high-quality information regarding recruitment, gear selectivity, growth, natural mortality, and relative depletion \citep{magnusson2007, ono2015}.
Furthermore, age and length samples are notoriously clustered \citep{hilbornwalters1992, francis2011}---fish of a similar age tend to be caught together---and removing spatial blocks from the survey domain could have a disproportionate impact on age composition data if the MPA happened to protect some non-random portion of the population (e.g., a nursery area for young fish, or a refuge or feeding area for older fish).
Finally, while spatiotemporal models are increasingly used for overall biomass or abundance density modelling and could use spatial information, covariates, or other surveys to alleviate the impact to indices, such approaches are less well-established for composition data \citep[but see][]{thorsonhaltuch2018}.

\subsection*{Caveats and next steps}

There are several caveats to our analysis.
First, we performed a retrospective analysis on a historical dataset.
A simulation approach \citep[e.g.,][]{schnute2003, regular2020, yalcin2023} could explore additional dimensions of the problem (e.g., systematically varying observation error, sampling intensity, or species distribution characteristics), derive potentially more generalizable results, and have a known truth for comparison instead of using the status quo index as a reference.
Conversely, using a historical dataset embeds the ``messiness'' of reality (e.g., poorly understood outliers, clustering, movement, and spatially varying abundance trends) and makes the results pertinent to real stocks and real surveys where management decisions are made.
% TODO: next is kind of duplicated elsewhere?
Second, we suspect that surveys covering larger areas (e.g., Gulf of Alaska) might fare better under survey restriction when using a spatiotemporal modelling approach compared to the smaller surveys examined here (especially West Coast Haida Gwaii).
Compared to other surveys, West Coast Haida Gwaii spatiotemporal indices were more affected by survey exclusion than their design-based equivalent.
This might be because the survey is on the verge of lacking sufficient spatial data to support these spatiotemporal models.
Third, we found that indices were most impacted for species that had a higher proportion of abundance or biomass within the MPAs.
These species might also be expected to be most protected by the MPAs and it is possible this protection could offset some of the risk incurred by poorer survey coverage.
However, the degree of protection will depend on several factors including juvenile and adult species movement \citep[e.g.,][]{gerber2003, gruss2011}, larval dispersal distance \citep[e.g.,][]{botsford2003, planes2009}, and \R{B4} enforcement of fishing restrictions.
\R{E5} Furthermore, knowledge about the effectiveness of the protection, and therefore developing effective species management strategies, will be challenging to establish without continued monitoring within the MPAs.

% TODO: come back to---duplicating elsewhere
% Our analysis of impacts on survey indices raises several important follow-on questions.
% We suspect that losing age and length composition samples may have a stronger effect on stock assessment estimates than the degraded survey indices examined here---such composition data are known to be important (REFs).
% Furthermore, added bias or imprecision on stock assessment estimates such as measures of status with respect to reference points and fishing intensity are only impactful if they affect science advice and the probability of achieving conservation and fishery objectives.
% Management strategy evaluation would be a key tool for addressing such questions.

% TODO: this paragraph is a bit rough still:

\R{B3}Omitting a portion of a stock from the survey domain results in a mismatch between the population spatial domain and the assessment domain \citep{field2006} and a violation of the unit stock concept---a basic theoretical assumption of nearly all conventional stock assessment \citep{russell1931, cushing1968}---specifically the assumption that there is no fish movement into or out of the stock spatial domain \citep[e.g.,][]{cadrin2020}.
Such a situation can lead to misperception of stock productivity and overharvesting or foregone yield \citep[e.g.,][]{goethel2011, mcgilliard2015, kerr2017}.
Furthermore, while there are a variety of approaches available to navigate the complexities of mismatched population and stock units \citep{kerr2017}, most would be challenged by a lack of monitoring data within the spatial domain of the MPA portion of a stock.
The degree this mismatch will impact stock assessment will depend on the magnitude of exclusion but also on the degree to which the fish move (i.e., connectivity) \citep{field2006}.
At one extreme, if the stock is stationary and has little larval exchange, the portion of the population in the MPA could be considered an independent stock and shrinking the survey domain would have little impact on assessment outcomes.
However, if a species freely moves into and out of the MPA---and especially if the stock does not maintain a static distribution---the survey will no longer represent an index that is proportional to abundance.
We expect most of the groundfish examined here fall towards the latter scenario.
Movement tracking, oceanographic studies of larval dispersal, and some degree of continued monitoring within the MPA could help inform these scenarios.

% If no longer represents the full population, spatial mismatch...

% Mismatch between population spatial domain and assessment domain or understanding of connectivity can ... \citep{kerr2017}

% \citep{mcgilliard2015}:
% e.g. of using spatial stock assessment to reduce impacts of MPAs on assessments
% "assessing populations as a single stock with use of fishery catch-rate data and without accounting for the NTMR results in severe underestimation of biomass for two of the movement pattern"

% Mismatch between population and assessment population structure (due to missing portions of sampling) can lead to biased stock status estimates, overharvesting, or foregone yield \citep{goethel2011}.

% "
% Development of assessment models that align with our understanding of biological structure can provide a more realistic assessment of population status and trends (Kerr and Goethel, 2014). Spatially-explicit models can account for connectivity within and across stock boundaries, but these models can be data intensive, add increased costs to monitoring (e.g. collection of tagging data for tag-integrated models), and require extensive knowledge of population structure in order to provide unbiased estimates (Porch et al., 1998, 2001; Goethel et al., 2015a,b)."

% "that specific stocks can be exploited independently of each other (unit stock assumption, Secor, 2014) or that catches can be assigned to their stock of origin (Cadrin et al., 2014a). Violation of the unit stock assumption (i.e. mis-classification of the appropriate spatial scale of management)"

% "Additionally, management units that are only a portion of a selfsustaining population can also pose problems for assessment and management due to difficulty in estimating biomass and providing catch advice for a portion of a population which may vary in its representation in a stock area over time (e.g. transboundary species that exhibit connectivity between United States and Canadian waters such as winter skate, Frisk et al., 2008; and Atlantic halibut, Shackell et al.,2016). Thus, understanding the spatiotemporal scale of population structure for a species in relation to management units is important for effective long-term sustainable management (Goethel et al., 2011)."

% "Our synthesis of the six case studies from the North Atlantic indicates that management bodies are becoming more aware of the importance of spatial structure and connectivity,"
% "The degree of spatial isolation or overlap between populations and harvest stocks are important determinants of the appropriate strategy."

% "“Simulations also illustrated that applying a harvest level based on misdiagnosed spatial structure leads to biased stock status indicators, overharvesting, or foregone yield. Across the scenarios examined, incorrectly specifying the connectivity dynamics (particularly misdiagnosing source–sink dynamics) was often more detrimental than ignoring spatial structure altogether.” ([Goethel and Berger, 2017, p. 1878](zotero://select/library/items/SC5B5PG8)) ([pdf](zotero://open-pdf/library/items/8IZV2AEA?page=1&annotation=YXCNTWDS))"

\subsection*{Recommendations}

If surveys are excluded from a portion of the original domain, our results suggest that survey effort should likely be redistributed, but this is not a panacea.
This will recover most of the lost precision, which is primarily lost due to reduced sample size.
However, the increased sampling outside the MPAs could make accuracy (with respect to the status quo index) worse, and do little to remedy biased perceptions of trend.
Our results suggest that impacts on trend can be caused by spatially contiguous survey loss combined with changes in species density within vs.\ outside the MPAs \citep[a form of spatially varying population trends,][]{barnett2021a}---no amount of additional sampling outside of the MPA can solve this problem.
In these cases, increased sampling merely makes us more precise about the ``wrong'' index because of the missing part of the population.

\R{E1}This missing part of the population also requires us to consider whether
we should be extrapolating into the excluded area. Recent simulation work has
shown that spatiotemporal models have some capacity to extrapolate into an
excluded area and recover the full survey-domain index, but the success of this
approach is heavily dependent on model configuration and a degree of MPA
spillover \citep{yalcin2023}. In our analysis, we found that extrapolation into
the excluded survey domain had limited benefit to accuracy but reduced
precision. However, even if the approach appeared promising with our
retrospective analysis, we caution against the approach as a long term solution
\R{A1} since covariate relationships might break down over time, distribution
driven by unmodelled biological and environmental variables (e.g., water
temperature, prey distributions) could change, and extrapolation will not
account for any changes in population dynamics outside vs.\ inside the MPA
beyond what can be observed with spillover. It is possible a reduced sampling
frequency within the MPA could could support continuing to predict across the
MPA and avoid the need for complete extrapolation.

Should we use design- or model-based indices when faced with survey restrictions?
Overall, we found similar effects for design- and model-based indices.
However, we see several advantages of the spatiotemporal model approach in this context.
First, the models produced more precise indices in the status quo scenario \citep[sensu][]{thorson2015a}, and even though proportionally they were more impacted by restriction than the design-based index, they remained more precise, on average.
% TODO: some duplication with elsewhere here:
Second, spatiotemporal model indexes are likely to be less impacted for surveys covering larger spatial areas or with denser spatial coverage than some surveys considered here.
We suspect the spatiotemporal models were already struggling in West Coast Haida Gwaii where large spatial chunks were then removed.
If spatial information, covariate effects \citep{johnson2019}, or even spatially varying covariate effects \citep{hastie1993, barnett2021a, thorson2023svc} could be estimated well, we suspect a spatiotemporal model may be able to improve accuracy compared to omitting the MPA from the survey domain prediction grid by extrapolating into the closed area, but at the expense of precision.
% TODO: some duplication with the previous point here:
Third, the model-based approach affords the opportunity to leverage other data sources such as other surveys \citep[e.g.,][]{gruss2019, webster2020, monnahan2021} to enhance indices.
The approach provides a mechanism to make use of historical data in areas that later restrict surveys rather than discarding all historical data the does not match new stratum definitions.
Fourth, the model-based approach provides a statistical framework to make inference about the benefits of the MPAs through causal inference \citep[e.g.,][]{pearl2016}, assuming some sampling within the MPAs is maintained.

Our results suggest that we should consider maintaining at least some survey effort within MPAs---using minimally destructive gear where possible.
Longline surveys will tend to have less benthic impacts than bottom trawl surveys; however, longline gear (1) will tend to select for larger bait-attracted individuals than caught in trawl gear and (2) tend to not sample the same species as trawl surveys.
Most groundfish species are well indexed by either longline or bottom trawl gear, and few species are indexed well by both \citep[e.g.,][]{anderson2019synopsis}.
Longline surveys are not, therefore, a direct substitute for trawl surveys in most cases.
% \citep{field2006}
We suggest accelerating research on less-destructive survey methods, such as underwater camera systems
% CR: ROVs are actually pretty uncommon in the survey literature. There's 1 ongoing ROV survey in the GOA for yelloweye rockfish I can think of, but most attempts at using cameras to survey fish use less expensive options (BRUVs, towed cameras, stationary cameras, etc). and of course manned submersibles, so a more general image based survey method statement might be better
\citep[e.g.,][]{trenkel2004, rooper2012, bryan2023} and eDNA approaches \citep[e.g.,][]{rourke2022, he2023}, but acknowledge these approaches are not a \R{B18}substitute for trawl surveys \citep{benoit2020national}.
% CR: Kind of disagree with this statement. These approaches will never fully replace trawl surveys in that they won't be able to collect specimen data. However, they are not that far from replacing trawl surveys in terms of estimating biomass. Especially acoustic surveys have gotten really good at replacing trawl surveys (or at least reducing trawling). I've got a bit better citation for this, I'll send it along
% SA: agreed, good points, text has been modified
In addition to changes in catchability and size selectivity, most alternative approaches lack an opportunity to collect biological samples.
Fish lengths may be more readily collected, such as visually with image-based surveys, but due to asymptotic growth, length provides little information on population age structure for a large portion of most groundfish lifespans.
% For example, a XX cm long TODO Rockfish could be XX or XX years old.
If new survey methods are adopted, it will be critical to calibrate new methods with historical ones to ensure unbroken time series \citep{field2006}.
As an alternative or supplement to alternative methods, some level of reduced survey intensity with extractive gear within closed areas might be chosen to balance benthic protection objectives with monitoring of other species groups.

\subsection*{Broader implications}

A reduction in the spatial domain of surveys and the degradation of index precision, accuracy, and trend estimation will impact our ability to make valid inference for other purposes beyond stock assessment.
Surveys support conservation planning analysis of biodiversity patterns and habitat suitability \citep[e.g.,][]{rubidge2016, thompson2022meps} and understanding species' potential responses to climate change \citep[e.g.,][]{english2021, thompson2023warming}.
Indices are also used by conservation agencies such as IUCN globally \citep{iucn2012} and COSEWIC in Canada \citep{cosewicTable2} to assess rates of decline and assign population status; the results of our power analysis to detect decline is particularly relevant to this use.
Surveys could also be used to monitor recovery effects of MPAs themselves against long historical baselines of expected spatial and temporal patterns and levels of observation error.
In fact, the surveys examined here were used to inform the MPA network design whose impacts we are evaluating.

% \textbf{Paragraph: Consider the risk tradeoff explicitly}

Eliminating scientific surveys inside protected areas shifts risk from those associated with localized disturbance of benthic habitats and organisms (including corals and sponges) onto a suite of commercially exploited species (groundfish here) by reducing the ability to effectively monitor their populations.
% If restricted fishing is not adequately enforced,
Many of these these commercially exploited species regularly move into and out of the MPAs and some are at low fractions of their unfished biomass.
How to balance this tradeoff is a policy decision, but we think it is important to remember that this tradeoff exists.
Decisions about how to manage this tradeoff should be informed by data analyses, such as what we have explored here, as well as estimates of survey gear recurrence time \citep{benoit2020national} and simulation modelling identifying impacts on probability of achieving conservation and fishery objectives.
% We suggest allowing at least some survey effort in closed areas, particularly from less destructive gear, accelerating research on less destructive survey methods, and ensuring sufficient calibration between any historical and new survey methods to maintain index integrity.

Although we applied our approach to a proposed MPA network, our approach is equally applicable to other forms of survey restriction including wind farms \citep{hare2022, haase2023}, oil and gas, funding cuts, or other logistical restraints \citep{ices2023}.
Structured decision making benefits from an understanding of consequence tradeoffs incurred by potential decisions \citep{gregory2012}.
Our approach is a structured way to assess expected impacts from unanticipated reductions in survey coverage using historical data.
% TODO: experimental? thinking adaptive management style experiments I guess
We see it as important step in analyzing the problem---complimented by simulation-based and experimental analyses.
% MPAs are an opportunity to monitor and learn about effectiveness of these strategies... but need to make sure they don't hamper our ability to
As long as we as a society let aquatic species be exploited commercially, we have an obligation to maintain the integrity of the long-term data streams that are critical to managing the health our marine ecosystems.
% Conservation efforts to support marine biodiversity should not, ironically, hamper our ability to assess marine species.

% \textbf{Paragraph: Do analyses such as these}
% We have unveiled a structured way to think about the problem
% Widely applicable to other surveys
% and impacts other than MPAs (e.g., wind/NOAA)
% Such be a first step in any decision making

% Longline surveys tend to have less benthic impacts than bottom trawl surveys, but longline doesn't sample all species well (e.g. POP) and selects for larger bait attracted individuals - in very few cases a substitute! \citep{field2006}
% accelerating research on less-destructive survey methods (ROV, eDNA, acoustic, ...), again, long way from being substitute \citep{field2006}
% but not just indices - also composition data and other life-history info (growth, maturity) - need some physical samples
% ensuring adequate calibration between historical and new less-destructive - includes length/age selectivity in addition to overall catchability \citep{field2006}
% It's possible a reduced survey intensity with destructive gear within closed areas could best balance benthic protection objectives with monitoring of other species groups.

% \citep{mcgilliard2015}:
% e.g. of using spatial stock assessment to reduce impacts of MPAs on assessments
% "assessing populations as a single stock with use of fishery catch-rate data and without accounting for the NTMR results in severe underestimation of biomass for two of the movement pattern"

% \citep{field2006}:
% "Although non-lethal means of estimating abundance (such as direct observation surveys from submersibles or towed vehicles) offer an alternative way to track abun- dance and size structure, particularly in areas difficult or impossible to sample with trawl or other extractive survey gear (Richards 1986; O’Connell and Carlile 1993; Jagielo et al. 2003)"
% Other non-lethal survey methods in- clude egg or larval abundance (or production) surveys (Mangel and Smith 1990; Moser et al. 2000; MacCall 2003; Ralston et al. 2003) and hydroacoustic surveys (Stanley et al. 2000; Helser et al. 2006),
%  However, any new indices of abundance based on non-lethal methods cannot replace historical time series without a period of overlap and calibration.

% Punt and Methot (2004) conducted a quantitative evaluation of the effects of implementing large-scale MPAs on the performance of stock assessment methods. Using the Management Strategy Evalua- tion approach,5 they found that while the negative impacts of MPAs on assessments were substantial if MPA and non-MPA data were aggregated, the impacts were minor when spatially structured models were used

% Bias and errors increased with movement rate, particularly when movement in- creased with age, because the models treated each area separately; performance was less sensitive to assumptions about larval dispersal and density dependence. Not surprisingly, a lack of fishery- independent survey data for the population inside the MPA led to considerably degraded model performance.

\section*{Acknowledgements}

We thank the many individuals who have contributed to collecting the survey data on which this manuscript is based.
We thank A.R. Kronlund, P.M. Regular, and S. Yalcin for helpful discussions.
We thank C.N. Rooper for comments on an earlier draft that improved this manuscript.
% TODO there will be more

\section*{Data Availability Statement}

The survey data used in this study are openly available through \href{open.canada.ca}{open.canada.ca} for the \href{https://open.canada.ca/data/en/dataset/524fde54-1d93-4d22-bb83-df542780a719}{HBLL OUT N longline survey} and the \href{https://open.canada.ca/data/en/dataset/a278d1af-d567-4964-a109-ae1e84cbd24a}{synoptic trawl surveys}.
Code for this analysis is available at \href{https://github.com/pbs-assess/gfmpa}{github.com/pbs-assess/gfmpa} and will be archived along with cached data through \href{https://zenodo.org/}{Zenodo} on publication.

\bibliography{refs}

\end{spacing}

\clearpage

% \section*{Tables}

% \clearpage

\section*{Figures}

\begin{figure}[htb]
  \centering
  \includegraphics[width=6in]{figs/fig1.png}
  \caption{Map of proposed Marine Protected Areas (MPAs) within the Northern Shelf Bioregion (NSB) MPA network that overlap with included surveys. Left panel shows the survey domains as shaded grey in 2 $\times$ 2 km grid cells and the MPAs coloured by zone category. Right panel shows all survey sampling locations between 2003 and 2022 as coloured dots (colour indicating survey) and with the overlapping MPAs coloured in black. Red square in the right inset indicates the area shown in the main panels. Acronyms: RCA (Rockfish Conservation Area), HBLL OUT N (Outside Northern Hard Bottom Longline survey), SYN HS (Synoptic Hecate Strait trawl survey), SYN QCS (Synoptic Queen Charlotte Sound trawl survey), SYN WCHG (Synoptic West Coast Haida Gwaii trawl survey). See the subsection ``Northern Shelf Bioregion zones'' and \citep{dfo2022networkactionplan} for a description of these zone categories. Category 1 and 2 zones are proposed for implementation by 2025 and 2030, respectively. Our analysis restricts longline survey access from Category 1, Category 2, Gwaii Haanas Strict + RCA zones. We restricted trawl survey access from these same zones plus Gwaii Haanas Strict.}
  \label{fig:map}
\end{figure}

% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=4in]{figs/map-existing-restrictions.png}
%     \caption{Map of current groundfish survey footprints within the Northern Shelf Bioregion. Survey colours ordered by increasing proportion covered by proposed MPAs (not illustrated) and dots represent all survey sampling locations between 2003 and 2021. Red polygons indicate existing restricted areas.}
%     \label{fig:map}
% \end{figure}

% 2. time series
\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/index-geo-restricted-highlights}
    \caption{Indices of relative abundance for selected species with variable patterns in abundance and responses to restrictions. Dots and lines represent means and 95\% CIs. Colours lines and points represent the status quo index (with colour indicating the survey) and grey represents the same time series estimated assuming survey sets within the NSB MPA network had been excluded and the survey domain shrunk. All time series have been centered to have a geometric mean of one and the upper vertical axis is limited to four times the maximum mean estimate per panel (affects Petrale Sole in 2007). See Figures \ref{fig:ts-hbll}, \ref{fig:ts-wchg}, \ref{fig:ts-hs}, and \ref{fig:ts-qcs} for all time series. ``Rougheye/Blackspotted'' refers to the Rougheye and Blackspotted Rockfish complex. Generalized additive models (GAMs) with gamma error, log links, and basis dimensions (k) of 8 are shown to highlight trends; note the flattening of some `restricted' trends due to increased interannual variation.}
    \label{fig:timeseries}
\end{figure}

% 3. CV/ MARE/ trend in RE
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.92\textwidth]{figs/metrics-dotplot-main}
    \caption{Variation across species in the loss of precision (a--d), accuracy (e--h), and trend bias in relative error (RE: i--l) for indices of abundance resulting from retrospective survey restriction in the shrunk-survey-domain scenario. Dots represent medians and lines represent interquartile ranges across annual index values (a--h) and the mean and 95\% CI on the estimated slope from a linear model of RE per decade (i--l). Vertical grey line indicates no change from the restriction and dashed coloured vertical lines indicate medians within a category (across all species, including those not shown). Surveys are sorted with those most impacted spatially by MPA restrictions at the top. Species are sorted from top to bottom within surveys by decreasing proportion of predicted biomass or abundance within the NSB MPA regions (shown in parentheses) and species are omitted if their proportions with the NSB MPA regions are less than 0.15 (1/26 for SYN WCHG, 0/19 for HBLL OUT N, 22/32 for SYN HS, 25/29 for SYN QCS; all species are shown in Fig.~\ref{fig:covariates}). See Fig.~\ref{fig:metrics-design-geo} for a version with design-based estimators and Fig.~\ref{fig:cv-abs} for absolute CVs. }
    \label{fig:dotplot}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{figs/metrics-cross-plot1.pdf}
  \caption{Relationships between status quo index uncertainty (CV), proportion of a species' biomass occurring within the NSB MPA network, and measures of accuracy and precision loss for indices of abundance in the shrunk-survey-domain scenario. Dots represent median values across years for individual species-survey combinations. Lines and shaded ribbons represent means and 95\% CIs from generalized linear models with gamma error and log links (a, b) or from a linear model with a non-constant variance (c). Colours represent individual surveys and the black line represents all surveys combined.}
  \label{fig:covariates}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=4.8in]{figs/slopes-wchg2}
    \caption{Slope in relative error (i.e., how ``wrong'' the estimated index trend is) as a function of trend in the proportion of a species' density that is located within the MPA network (i.e., whether a species is increasing in density inside vs.\ outside the MPA). Shown here is the SYN WCHG survey where the strongest trends occurred. Species to the right of zero are increasingly hidden by the MPAs; species to the left, proportionally, are increasingly found outside the MPAs. Panels on the right illustrate the most extreme cases of trends in density proportion within the MPAs with GAMs fit for visualization. Colours in left and right panels match. Solid sloped line and ribbon represent a regression mean and 95\% CI; dotted sloped line represents a 1:1 relationship.}
    \label{fig:slopes}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/sampled-dotplot-comparison2}
    \caption{Evaluating the contribution of reduced sampling effort. Random down-sampled is a scenario where we did not remove survey sets from the MPAs explicitly, but randomly (within strata) removed an equivalent level of survey effort from the entire survey domain. This evaluates the role of the particular location and blocked nature of the MPA removals vs.\ survey effort loss alone. Random-up-sampled and shrunk (blue) is a scenario where an equivalent number of survey sets lost in the MPA network are simulated in the remaining survey domain to represent a redistribution of the same effort. Restricted and shrunk is the standard MPA-restricted scenario illustrated elsewhere in the analysis. Panels a, c, and d show the same metrics as in Fig.~\ref{fig:dotplot}. Panel b shows the proportion of replicates across species (and five random seeds) in which the 50\% confidence intervals include the status quo index mean.}
    \label{fig:sampling}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=3.0in]{figs/power-fig7}
    \caption{Statistical power to detect a simulated population decline of a given magnitude over the span of the survey (15 [HBLL OUT N], 16 [SYN WCHG, SYN HS], or 18 [SYN QCS] years). Arrows begin at power under the status quo scenario and end at power under the scenario where surveys are restricted from the MPA network; longer arrows represent greater decreases in power. We explored two levels of population decline depending on the status quo survey CV: 50\% decline for surveys with CVs $>$ 0.25 and 30\% decline for surveys with CVs $\le$ 0.25. We fit models with spatial random fields and a linear fixed effect of year (in log link space) on abundance or biomass density. We then defined power as the proportion of 300 iterations that the upper 95\% CI on year fixed effect was correctly below zero. Species are arranged by length of arrows within decline categories. Numbers in parentheses represent average status quo index CVs. This version only shows species with $>$ 15\% of density within the MPA network; see Fig.~\ref{fig:power-supp} for all species.}
    \label{fig:power}
\end{figure}

% % 6. trend in MPA vs. trend in RE
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.8\textwidth]{figs/explore-all-slopes2.pdf}
%     \caption{Relationship between the proportion of a species' biomass occurring within MPAs and the change in relative error per decade (RE trend) resulting from `extrapolated' indices of abundance from all four surveys. Selected species are labelled.}
%     %  TODO: Maybe update fig 2 to include Silvergray and others highlighted for QCS and HBLL here?
%     \label{fig:slopes}
% \end{figure}

% - HBLL N and QCS

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{section}{0}
\setcounter{subsection}{0}
\setcounter{subsubsection}{0}

\clearpage

\appendix
% \internallinenumbers

\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}

\setcounter{secnumdepth}{0} % Removes section numbers

\onehalfspacing
\linenumbers
\resetlinenumber
\setcounter{page}{1}
\setcounter{equation}{0}
\nolinenumbers

\begin{Center}
\section*{Supporting Information}
\end{Center}

\section*{Supporting Methods}

\subsection*{Spatiotemporal modelling}

We employed the stochastic partial differential equation (SPDE) approximation to Gaussian random fields for computational efficiency and fit our models with the \textsf{R} package sdmTMB \citep{anderson2022}, which uses INLA \citep{lindgren2015} for SPDE ``mesh'' construction, TMB \citep{kristensen2016} to calculate the marginal log likelihood, and the \texttt{nlminb} non-linear optimizer in R \citep{r2023} to find the parameter values that maximize the marginal log likelihood.
We constructed the meshes with vertices at least 10 km apart.
We ensured the optimization of all models was consistent with convergence by checking that the Hessian was positive definite, the maximum gradient with respect to fixed effects was $<$ 0.001, and all random field marginal standard deviations were $\ge$ 0.01.
From our fitted models, we calculated area-weighted indices of abundance by predicting on a 2 $\times$ 2 km grid covering the survey domain, summing the predicted biomass or abundance each year \citep{thorson2015a}, and applying a generic bias-adjustment given the non-linear transformation of the random effects \citep{thorson2016bias}.

Our models took the general form:
\begin{align}
\mathbb{E}[y_{\boldsymbol{s},t}] &= \mu_{\boldsymbol{s},t},\\
\mu_{\boldsymbol{s},t} &=
f^{-1} \left(
\alpha_g +
O_{\boldsymbol{s},t} +
\omega_{\boldsymbol{s}} +
\epsilon_{\boldsymbol{s},t} \right),
\end{align}
\noindent
where
$y_{\boldsymbol{s},t}$ represents the response data at point $\boldsymbol{s}$ and time $t$ (catch weight or count),
$\mu$ represents the mean,
$f$ represents a link function (here logit or log) and $f^{-1}$ represents its inverse,
$\alpha_{g}$ represents independent intercepts for each year,
$O_{\boldsymbol{s},t}$ represents an offset (log transformed area swept or hook
count),
$\omega_{\boldsymbol{s}}$ represents a spatial random field with Mat\'ern covariance [$\omega_{\boldsymbol{s}} \sim \mathrm{MVN}(\boldsymbol{0},\boldsymbol{\Sigma}_\omega)$],
and $\epsilon_{\boldsymbol{s},t}$ represents spatiotemporal random fields with Mat\'ern covariance [$\epsilon_{\boldsymbol{s},t} \sim \mathrm{MVN}(\boldsymbol{0},\boldsymbol{\Sigma}_{\epsilon})$].
The delta-gamma models have two components: a binomial component with a logit link modelling species encounter vs.\ non-encounter and a gamma component with a log link modelling species catch (biomass or count) if a species was observed in a given set.
The Tweedie models use \citet{tweedie1984} observation error with a log link.

We place penalized complexity priors \citep{fuglstad2019} on the Mat\'ern covariance parameters. This weakly constrains the parameters to reasonable values and will tend to favour simpler random fields (less `wiggly' fields with smaller standard deviations and larger ranges) if the data are not informative about spatial correlation. We set priors on both the spatial and spatiotemporal marginal standard deviation as $P(\sigma > 5) = 0.05$ and a prior on the range parameter shared across any spatial and spatiotemporal fields within a model component as $P(\rho < 20\ \mathrm{km}) = 0.05$.

We fit a sequence of models to emulate what an applied modeller might do if manually fitting models to these data sets and choosing a model that is best able to represent variation in the data while achieving satisfactory convergence diagnostics (Table~\ref{tab:model-configs}). For each survey and species combination and in the restricted survey scenario, the first model is fit; if that model fails convergence checks (positive-definite Hessian matrix, maximum gradient with respect to fixed effects $<$ 0.01, all random field standard deviations $\ge$ 0.01) then the subsequent model is run and so on. We start by fitting models to the ``restricted'' survey data sets since these have the least data and therefore tend to have the most convergence issues. We then record the chosen model and use this same model when fitting the full data set (and any up- or down-sampled datasets) for comparison (Fig.~\ref{fig:metrics-by-model}).

\subsection*{Northern Shelf Bioregion zones}

The Northern Shelf Bioregion MPA network proposes `Category 1' and `Category 2' zones with the main distinction that Category 1 zones are proposed for implementation by 2025 and Category 2 zones by 2030 \citep{dfo2022networkactionplan}.
In our analysis, we assume both zones are restricted from bottom trawl and long-line research surveys.
From the existing Gwaii Haanas National Marine Conservation Area Reserve and Haida Heritage Site, we excluded trawl surveys from all ``Strict Protection'' zones, while only excluding the long-line survey from ``Strict Protection'' zones that overlap with Rockfish Conservation Areas. We do not exclude survey activity from remaining existing MPA zones where possible changes to activities are uncertain at this time \citep{dfo2022networkactionplan}.

% CR: Similar comment for this paragraph, to the casual reader (e.g. not from BC), these details may not mean much.
% SA: I think it's important to keep; was important to ER etc. plus this it the 'case-study' section
Among the Category 1 and 2 zones, our analysis represents a worst-case scenario for survey indices, since it is unlikely that all survey effort would be restricted from all areas.
In reality, access to sites for research may vary depending on the area's designation and the potential for the surveys to contribute to monitoring the MPA.
We also assume all Category 1 and 2 zones will ultimately be implemented as MPAs.

\clearpage

\section*{Supporting Tables}

\input{figs/spp-table.tex}

\clearpage

\begin{table}[htpb]
\caption{Sequence of model configurations attempted. All configurations have
spatial random fields. The final model chosen for each species-survey combination is shown in Fig.~\ref{fig:metrics-by-model}.}
\centering
\begin{tabular}{lllll}
\toprule
Model run & Family & Spatiotemporal fields  \\
\midrule
1 & Delta-gamma & Included in binomial and gamma components  \\
2 & Delta-gamma & Included in gamma model only \\
3 & Tweedie & Included  \\
4 & Delta-gamma & None \\
\bottomrule
\end{tabular}
\label{tab:model-configs}
\end{table}

\clearpage

\section*{Supporting Figures}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/grid-strata-restricted.pdf}
    \caption{Depth strata (coloured cells) and their overlap with the NSB MPA network (black-outlined cells).}
    \label{fig:strata}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/ts-hbll.pdf}
    \caption{Outside Northern Hard Bottom Longline survey (HBLL OUT N) indices. Dots represent means and line segments represent 95\% confidence intervals. Colour represents the full ``Status quo'' survey and grey represents the survey if it had been restricted by the NSB MPAs. All time series have been centered to have a geometric mean of one to eliminate changes to the scale of the indices. Species are ordered by decreasing proportion of density within the NSB MPAs from top to bottom. Curved lines represent fitted generalized additive models with gamma error distributions and log links to visually highlight differences in longterm trend.}
    \label{fig:ts-hbll}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/ts-wchg.pdf}
    \caption{West Coast Haida Gwaii (SYN WCHG) indices. Caption otherwise the same as for
    Fig.~\ref{fig:ts-hbll}.}
    \label{fig:ts-wchg}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/ts-hs.pdf}
    \caption{Hecate
      Strait (SYN HS) indices. Caption otherwise the same as for Fig.~\ref{fig:ts-hbll}.}
    \label{fig:ts-hs}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/ts-qcs.pdf}
    \caption{Queen Charlotte Sound (SYN QCS) indices. Caption otherwise the same as for Fig.~\ref{fig:ts-hbll}.}
    \label{fig:ts-qcs}
\end{figure}

\clearpage

% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=\textwidth]{figs/ts-qcs-hs.pdf}
%     \caption{Queen Charlotte Sound (SYN QCS) and Hecate
%       Strait (SYN HS) indices. These two neighbouring
%   surveys that are conducted one after the other with the same gear have been
% combined into one index. Caption otherwise the same as for Fig.~\ref{fig:ts-hbll}.}
%     \label{fig:ts-qcs-hs}
% \end{figure}

% \clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.90\textwidth]{figs/abs-cv}
    \caption{Coefficient of variation (CV) of design-based and spatiotemporal-model-based indices of abundance. Whereas Fig.~\ref{fig:dotplot}a--c shows percent increase in CV if the survey was restricted from MPAs, this figure shows the absolute CV. Dots show medians and line segments show the interquartile range (across years). Dashed horizontal lines show the overall median within a method and survey for the illustrated species. Only species with at least 15\% of their biomass or abundance density within the MPA region are shown (numbers in parentheses beside species names). Vertical black lines show the stated target of CV = 0.2.}
    \label{fig:cv-abs}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.99\textwidth]{figs/metrics-dotplot-extrapolate}
    \caption{Same as Fig.~\ref{fig:dotplot} but adding a version where the species density is extrapolated into the restricted areas instead of reducing the survey domain and showing all species.}
    \label{fig:dotplot-extrapolate}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.99\textwidth]{figs/metrics-dotplot-design-geo}
    \caption{Same as Fig.~\ref{fig:dotplot} but adding a version where the indices of abundance were calculated as random stratified design-based indices.}
    \label{fig:metrics-design-geo}
\end{figure}

\clearpage

% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.8\textwidth]{figs/metrics-dotplot-main-design}
%     \caption{Same as Fig.~\ref{fig:dotplot} but with design-based indices for all species where a design-based estimator could be calculated. Unlike Fig.~\ref{fig:metrics-design-geo}, this version includes some rarer species for which geostatistical models could not be fit.}
%     \label{fig:metrics-design}
% \end{figure}

% \clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=5in]{figs/prop-mpa-vs-metrics-design2.pdf}
    \caption{Precision loss, accuracy loss, and trend bias as a function of proportion of stock in the MPA for the scenario where survey effort is restricted in the MPAs and the survey domain is shrunk. This is similar to Fig.~\ref{fig:covariates} but (1) adds design-based indices (left column) and (2) extends the left side of the x-axis to include species with low proportions within the MPA. Colours represent individual surveys and the black lines represent all surveys combined.}
    \label{fig:covariates-design}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figs/geostat-vs-design2}
    \caption{Model- vs.\ design-based CVs. Dashed line shows the one-to-one line. The spatiotemporal-model-based indices tend to have higher precision (lower CVs) than the design-based bootstrap estimators. Panel a shows the status quo scenario and panel b shows the scenario where survey effort has been excluded from the MPA network.}
    \label{fig:design-model-cv}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.95\textwidth]{figs/metrics-dotplot-by-model.pdf}
    \caption{Illustration of which spatiotemporal model was selected for each species-survey combination. ``Off'' means no spatiotemporal random field was used for that component. For example, ``Off, IID'' means the binomial component had no spatiotemporal random effects but the gamma component has spatiotemporal random effects that were independent each year (but with a common standard deviation). All models included spatial random effects and independent means for each year.}
    \label{fig:metrics-by-model}
\end{figure}

\clearpage

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.95\textwidth]{figs/power-supp}
    \caption{Statistical power to detect a simulated population decline of a given magnitude over the span of the survey. Caption is the same as Fig.~\ref{fig:power}; however, this version shows all species and survey combinations whereas Fig.~\ref{fig:power} only shows species with greater than 15\% of biomass or abundant density within the MPA network.}
    \label{fig:power-supp}
\end{figure}

\clearpage

\section*{Response to reviewer comments}

\setcounter{page}{1}
\nolinenumbers

\setlength{\parskip}{5pt}%
\setlength{\parindent}{0pt}%

\section{Editor:}

\rev{E1: Line 114 “We calculated a series of indices. The first two form the main analysis and the third is an additional option we focus on in supplemental figures.” – I suggest that in discussion you spend some time on it. Some explicit advice to readers on what approach to use and when will be useful.}

We have added a paragraph discussing these approaches (see \lr{E1}).

\rev{E2: line 318 “Furthermore, biases in population trend—driven largely by changes in distribution inside vs. outside the excluded areas—could impact stock assessment results and estimates of decline for conservation agencies.” In my mind ability to estimate decline (or increase ) for conservation agencies should be consider as very important. If the role of MPAs is to protect something we need to know how well this protection is working. Eliminating ability to monitor stock in MPA can be counterproductive for conservation purposes. I suggest that you develop this topic a little bit more.}

TODO MORE HERE!! We agree. We have added text on this topic in response to Reviewer 2's comment `B3'. See the new text starting at \lr{B3}.

\rev{E3: Line 344 “changes to catchability due to different fishing captains” Changes in catchability are caused by different captains, but also and probably to larger degree by other factors such as variation in spatial and vertical availability to surveys, weather and environmental conditions (e.g. light, temperature), use of multiple nets, fish density, etc. The additional variance component in the assessment would increase if the MPA are excluded from surveys. I suggest that you expand on this a little bit.}

Thank you for highlighting this. We have added to the list of factors affecting catchability on \lr{E3a} and added a sentence about adding to this additional process error on \lr{E3b}.

\rev{E4: Line 368 “length composition data.” – add “and age”}

This now reads ``age and length composition data". See \lr{E4}.

\rev{E5: Line 400 “These species might also be expected to be most protected by the MPAs and it is possible this protection could offset some of the risk incurred by poorer survey coverage.” But it is also possible that lack of adequate monitoring can be detrimental to management of these species. When taking protective action is also necessary to monitor effectiveness of these actions.}

We agree. We think this is the main point of the paper, but here we were adding one possible caveat. We have revised the final sentence in the paragraph to expand on this a bit (see \lr{E5}): ``Furthermore, knowledge about the effectiveness of the protection, and therefore developing effective species management strategies, will be challenging to establish without continued monitoring within the MPAs.''

\section{Reviewer: 1}

\rev{A1: Overall the paper is very well written. The discussion section, in particular, does a great job of providing interpretation of results and considers the majority of potential future scenarios or assessments where their advice might change. There are many other considerations to this analysis, but the authors do a good job capturing of what most of those future studies are that are needed.}

\rev{General comments:}

\rev{A1: -include the idea in the discussion that this all assumes there are no shifts in distribution in response to ongoing environmental changes (e.g., water temps, predator/prey compositions and spread). I think it’s important to stress that in addition to closures and changing fishing pressure, there are other changes ongoing that can also be simultaneously impacting how the existing surveys capture the populations, and cutting out portions of that will make it even more difficult to disentangle all of this.}

We now emphasize this in the new paragraph starting on \lr{E1}. In particular, see \lr{A1}.

\rev{A2: Line 68: I would be specific on how this decision can impact scientific advice to management (E.g., TAC will change) since this is the point that provides management relevance to the research. “This decision can also impact index quality and therefore scientific advice to management, for example by”…}

We have added some text to this sentence. See \lr{A2}.

\rev{A3: Lines 160 – 170: Glad to see the down-sampled but not explicitly from the MPA domain…. Helpful to have a noise scenario for comparison.}

Thanks, we agree.

\rev{A4: Line 182: what is ``meaningful'' defined as regarding the rates of decline chosen? Need a little more info on that.}

We have added ``in a conservation context'' (see \lr{A4}). The specifics of how we defined that are in the latter part of the sentence: ``50\% decline for species with average status quo CVs of $>$ 0.25 and 30\% decline for species with CVs $\le$ 0.25.''.

\rev{Figures:}

\rev{A5: Figure 1. Can you provide the definitions/scenarios for each of the acronyms in the panels right in the figure legend?}

We have added definitions for all of the acronyms, described the red square in the inset, and described the restrictions we assumed with various zone categories. See the Figure 1 caption.

\rev{A6: Figure 5. Really nice figure.}

Thanks!

\section{Reviewer: 2}

\rev{Overall comments}

\rev{B1: This is an important and timely paper. Anderson et al. carefully analyze the impact of exclusions from MPAs on survey indices. As raison d’eˆtre, I fully agree with the last statement of the Abstract “Regulating surveys within MPA boundaries therefore requires careful consideration to balance MPA objectives with the need for reliable data.” The subject matter is highly important in many regions (we have 30\% by 2030 in parts of Europe also).}

Thanks!

\rev{B2: I like the approach of retrospectively analysing the real data as if an MPA were enacted. This is careful and targetted and the authors do a good job of the comparison to a simulation approach in the Discussion. Often, simulations can be very difficult to do right and their detail detract from the clarity of the message. It is also very useful to have defined performance measures: precision, inter-annual accuracy, and bias in indices derived from these surveys.}

Thank you, we agree.

\rev{My comments are largely minor as the paper is exceptionally well written and sculpted.}

\rev{B3: I do worry about the impact on the stock assessment (and reference points) and agree with the authors that transfer in and out of the MPA is likely. The authors do a nice job of discussing this within the confines of their results. I guess that if the MPA is a source, the assessment believes that the stock in the sampled area is more resilient than it is. Better to continually sample within the MPA and build this in. Maybe link to some of Lisa Kerr’s (or others) work on spatial assessments in the Discussion?}

Good suggestion. We now bring in literature on spatial stock assessment and population/stock scale mismatch. See the paragraph starting on \lr{B3}.

\rev{B4: Enforcement is a challenge for MPAs in many regions globally (massively so in Europe). If some sectors break the rules of entering MPAs, and others---charged with monitoring---remain outside it could be a double-whammy of no monitoring in areas with unknown exploitation.}

This is a good point. We mention this now on \lr{B4}. We couldn't find a place to give this more space without straying too far from the analysis.

\rev{Minor comments}

\rev{[] suggest delete () suggest include}

\rev{Abstract}

\rev{B5: Survey exclusion [especially affected] (particularly impacted) less precise indices,}

Done

\rev{Keywords}

\rev{B6: A Google Ngram suggests the primacy of “spatiotemporal” now. Probably time to
fully concatenate the adjective!}

Interesting. We removed ``spatio-temporal''.

\rev{Methods}

\rev{B7: Line 121 “although in the future some survey effort may be redirected over the
shrunk survey domain”, is this a complication worth mentioning here?
I immediately wanted to know when and how much!}

Good point. Removed.

\rev{B8: Under the correlates: “For any species with slight proportional decreases in precision, we set these values to small positive values (0.01) for the purposes of this GLM because the log link requires positive data.”}

\rev{B9: This seems a bit of a corner to cut. Would a normal distribution with non-constant variance fit it and avoid the fudge?}

Fair point. Looking back, we made this decision when a small number of species-survey combinations were just slightly below zero. In the final version, there are some outliers well below zero. We have fit a linear model here that includes a non-constant variance (a linear of effect of proportion of density in MPA on the log dispersion variance). This affects Figure 4 and Figure S9.

\rev{B10: For the “HBLL OUT N” survey, are all line sets deployed for the same amount of
time or is there an offset to include here too?}

The aim is to deploy all line sets for the same duration (2 hours). There is a bit of variation, but typically when using these data in an applied setting, the soak time has been considered constant. There is an offset here for log hook count (see \lr{B10}).

\rev{B11: Line 174 “We focused our analysis on species with at least 15\% of their biomass or abundance density within the MPA network and only for the surveys that were most impacted in precision and accuracy” I'm not convinced I would restrict it to those surveys most impacted. Would this give a biased result of the loss of power as it focusses on those surveys most impacted?}

Fair point. We have now run all species and survey combinations that could be estimated. We added the remaining two surveys to Fig.~\ref{fig:power}. We still restrict the main figure to species with $>$ 15\% of their density within the MPA network (same as Fig.~3) to save space but include a new supplementary figure that includes all species Fig.~\ref{fig:power-supp}. At the same time, we increased the number of simulation replicates (now 300) to make sure the figure would be consistent across random seeds.

% This might have been poorly worded. We focused a main figure (Figure 3) on those species with at least 15\% of their density within the MPA network out of necessity to save space.The other figures include all species for which we could calculate an index and the overall dashed vertical lines in Figure 3 indeed include all species.
% We indeed focused on only those surveys impacted by the MPA network, but other related surveys are completely avoided by the MPA network making them a not very interesting element to include. We have clarified this. See \lr.

\rev{Results}

\rev{B12: Expectedly brilliant visualisation throughout.}

Thanks!

\rev{B13: Is there something up with the GAM fit to Yellowmouth Rockfish in Figure 2. The indices don’t abviously follow the GAM trend. I might have missed if the degrees of freedom were restricted in these GAM fits.}

There shouldn't be anything wrong with the GAM fit. They're all created in one call to ggplot. The dimension of the basis functions was limited to 8, and we added this to the Figure 2 caption. We think what's happening is the interannual variability in the index values can't be fit well by the GAM and so the smoother degrees of freedom are penalized resulting in a flat line in link space for Yellowmouth Rockfish. Then, since it's a log link and an increasing index, it must curve upwards in natural space.

\rev{B14: Rougheye/Blackspotted has larger CIs in Figure 2 but a decrease in precision in
Figure 3a?}

We think this was a simple mistake in reading the figures on the reviewer's part. The Rougheye/Blackspotted panel in Figure 2 is purple, which indicates it's for the SYN HS survey. The CV indeed increases in panel Figure 3\textbf{c}. Figure 3\textbf{a} is for the SYN WCHG survey.

\rev{B15: Figure 4 is highly informative for all, but particularly decision makers.}

Thanks!

\rev{B16: Figure S9 could be brought into the test of the Results (around Line 283) a little better as the impact of MPAs on MARE and RE trend less for geostatistical than design-based. Whereas its the opposite for the increase in CV. This is mentioned in the Discussion but could be included here.}

Good point. We have done this starting on \lr{B16}.

\rev{Discussion}

\rev{B17: "more precise about the "wrong" index" seems Abstract-worthy.}

Good idea. We have added this. See \lr{B17}.

\rev{B18: Line 458 "are not a complete substitute for trawl surveys" - the decomposition of multiple nuclear DNA copies from multiple individuals in a turbulent flow seems to render it a little ways off. Personally I’d delete ``complete'' here}

We removed "complete". See \lr{B18}.

\end{document}